Using Distributed Information Stores for Patient-Clinician Health Data Sharing

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex
Added refs

\section{Introduction}
A key goal in improving patient-clinician communication is
establishing a common context for communication with which a shared
understanding can be established.  This understanding is
bi-directional, in the sense that the patient needs to be able to
effectively interpret what the clincian is trying to convey to him or
her, but also that the clinician needs to be able to arrive at an
accurate characterisation of the patient's situation and condition, as
updated since the patient was last seen. This is often under time
pressure and with limited access to information about how the patient
has been in the interim.

One of the most promising technologies that has the potential to help
provide such a common context is the patient-centric electronic health
record (henceforth EHR)\cite{ehr}, a digital representation of the
complete medical history of an individual as recorded by clinicians
throughout a patient's life. The EHR has enabled both patients and
their clinicians unfettered access to the patient's medical data,
including diagnostics, reports, and history of interventions.  This
means that clinicians can more easily review a patient's long-term
history even under time pressure and constriants. Moreover, since
patients have access to their own medical histories, they can access
the raw information that was the basis for a prognosis or decision.

@@

Fortunately, many of these sites have APIs that provide access to the raw sensor data, making it feasible to access and interpret captured data by way of these services.  We propose that these APIs be used to consolidate information from among all of the services corresponding to a patient's personal health devices, into a single reprsentation kept with their EHR.  In particular, grounding this activity data in a common taxonomy with associated description logic, like those provided for medical diagnoses and diseases through SNOWMED-CT \cite{stearns2001snomed} and ICD-10 \cite{icd}, could enable these diaries to interpreted directly by clinical medical systems, such as primary care information management systems.

>>>>

\section{Introduction}
A key goal in improving patient-clinician communication is
establishing a common context for communication with which a shared
understanding can be established.  This understanding is
bi-directional, in the sense that the patient needs to be able to
effectively interpret what the clinician is trying to convey to him or
her, but also that the clinician needs to be able to arrive at an
accurate characterisation of the patient's situation and condition, as
updated since the patient was last seen. This is often under time
pressure and with limited access to information about how the patient
has been in the interim.

One of the most promising technologies that has the potential to help
provide such a common context is the patient-centric electronic health
record (henceforth EHR)\cite{hayrinen2008definition}, a digital representation of the
complete medical history of an individual as recorded by clinicians
throughout a patient's life. The EHR has enabled both patients and
their clinicians unfettered access to the patient's medical data,
including diagnostics, reports, and history of interventions.  This
means that clinicians can more easily review a patient's long-term
history even under time pressure and constraints. Moreover, since
patients have access to their own medical histories, they can access
the raw information that was the basis for a prognosis or decision.

@@

Fortunately, many of these sites have APIs that provide access to the raw sensor data, making it feasible to access and interpret captured data by way of these services.  We propose that these APIs be used to consolidate information from among all of the services corresponding to a patient's personal health devices, into a single reprsentation kept with their EHR.  In particular, grounding this activity data in a common taxonomy with associated description logic, like those provided for medical diagnoses and diseases through SNOWMED-CT \cite{stearns2001snomed} and ICD-10 \cite{world1993icd}, could enable these diaries to interpreted directly by clinical medical systems, such as primary care information management systems.
